---
title: "Creating the Manuscript Plots"
author: "Lucas A Salas, Samuel Reynolds"
output: html_document
date: "2024-01-12"
---

Loading required packages.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(cowplot)
library(pheatmap)
library(IDOL)
library(meffil)
library(tidyr)
```

Loading methylation libraries.

```{r}
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/Mouse IDOL Minfi Noob libraries.rda")
```

Reporting the performance of each model - RMSE and R^2.

```{r}
# 200 "The Average RMSE = 10.595 and R2 = 0.688 for the IDOL Optimized Library"
#250 "The Average RMSE = 10.436 and R2 = 0.759 for the IDOL Optimized Library"
#300 "The Average RMSE = 10.016 and R2 = 0.752 for the IDOL Optimized Library"
#350 "The Average RMSE = 9.915 and R2 = 0.756 for the IDOL Optimized Library"
#400 "The Average RMSE = 9.793 and R2 = 0.744 for the IDOL Optimized Library"☺
#450 "The Average RMSE = 10.188 and R2 = 0.76 for the IDOL Optimized Library"
#500 "The Average RMSE = 10.251 and R2 = 0.757 for the IDOL Optimized Library"
#550 "The Average RMSE = 10.317 and R2 = 0.756 for the IDOL Optimized Library"
#600 "The Average RMSE = 10.359 and R2 = 0.763 for the IDOL Optimized Library"
#650 "The Average RMSE = 10.785 and R2 = 0.768 for the IDOL Optimized Library"
#700 "The Average RMSE = 10.601 and R2 = 0.757 for the IDOL Optimized Library"
matrixEPICIDOL_All<-matrix(data = NA, nrow=11, ncol = 2,
                           dimnames = list(paste0("CpG_",
                                                  seq(from = 200, to = 700, by =50)),
                                           c("RMSE", "R2")))
for (i in  seq(from = 200, to = 700, by =50)){
  IDOLtmp<-get(paste0("IDOLopt_mouse_minfiNoob_",i))
  for (j in 1:500){
    if (j==1){
      RMSE=RMSE.null=IDOLtmp$RMSE[j]
      R2=R2.null=IDOLtmp$R2[j]
    }else{
      if (IDOLtmp$RMSE[j] <= RMSE.null & IDOLtmp$R2[j] >= R2.null) {
        RMSE.null = IDOLtmp$RMSE[j]
        R2.null = IDOLtmp$R2[j]
        #print(paste("Iteration: ", j, " RMSE=",
        #    round(IDOLtmp$RMSE[j], 3), "; R2=", round(IDOLtmp$R2[j], 3),
        #    sep = ""))
      }
    }
  }
  print(paste("The Average RMSE = ", round(RMSE.null,
                                           3), " and R2 = ", round(R2.null, 3), " for the IDOL Optimized Library ",i,
              sep = ""))
  matrixEPICIDOL_All[paste0("CpG_",i), 1]<-RMSE.null
  matrixEPICIDOL_All[paste0("CpG_",i), 2]<-R2.null
}
```

Creating the R^2 and RMSE plots.

```{r}
# R2 and RMSE Plots
matrixEPICIDOL_All <- as.data.frame(matrixEPICIDOL_All)
matrixEPICIDOL_All$`Library Size` <- seq(from = 200, to = 700, by =50)
matrixEPICIDOL_All$`Library Size` <- as.factor(matrixEPICIDOL_All$`Library Size`)

blue_palette <- c("#bdd7ef", "#a3c2e8", "#8aaedf", "#719ac8", "#5987b3", "#43759e", "#30638a", "#1e5176", "#0e405f", "#002f4a", "#001f34")

# R2 Plot
R2Plot <- ggplot(data = matrixEPICIDOL_All, aes(x = `Library Size`, y = R2, fill = `Library Size`)) +                 geom_bar(stat = "identity") +
  scale_fill_manual(values = blue_palette) +
  labs(x= "Deconvolution Library Size", y= "R2",) + coord_cartesian(ylim=c(0.83, 0.9)) +
  geom_text(aes(label = round(R2,3)), vjust = -0.25, size = 3, 
            fontface = "bold") +
  theme(legend.position = "none",
        axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left = element_line(color = 'black'),
        axis.line.y.right = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
R2Plot

# RMSE Plot
RMSEPlot <- ggplot(data = matrixEPICIDOL_All, aes(x = `Library Size`, y = RMSE, fill = `Library Size`)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = blue_palette) +
  geom_text(aes(label = round(RMSE,2)), vjust = -0.25, size = 3,
            fontface = "bold") + 
  labs(x= "Deconvolution Library Size", y= "RMSE",) + coord_cartesian(ylim=c(7.5, 10)) +
  theme(legend.position = "none",
        axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left = element_line(color = 'black'),
        axis.line.y.right = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
RMSEPlot

# Buliding a combined plot.
R2_and_RMSE_Plot <- plot_grid(R2Plot, RMSEPlot, nrow = 2, labels = c("A","B"))
R2_and_RMSE_Plot
ggsave(filename = "~/Desktop/Christensen Lab/Mouse Methylation/Final Figures/Final R2 and RMSE Plot.png", plot = R2_and_RMSE_Plot, units = "in", height = 8, width = 8)
```

Creating the clustered methylation heat maps across cell types.

```{r}
# Loading the beta values and phenotype data.
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/allBetas.RData")
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/allPheno.RData")

# IDOL DMCs
allBetas_DMCs <- allBetas[row.names(allBetas) %in% IDOLopt_mouse_minfiNoob_300$`IDOL Optimized Library`,]
allBetas_DMCs <- allBetas_DMCs[,colnames(allBetas_DMCs) %in% allPheno$ID]
allBetas_DMCs <- as.data.frame(allBetas_DMCs)

# Taking the average beta value across cell types.
allBetas_DMCs$B <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "B", c("ID")]])
allBetas_DMCs$CD4 <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "CD4", c("ID")]])
allBetas_DMCs$CD8 <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "CD8", c("ID")]])
allBetas_DMCs$Mono <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "Monocyte", c("ID")]])
allBetas_DMCs$NK <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "NK", c("ID")]])
allBetas_DMCs$PMN <- rowMeans(allBetas_DMCs[, allPheno[allPheno$CellType == "PMN", c("ID")]])

# Creating the two data frames to be used in the heatmaps - averaged betas and individual samples.
allBetas_DMCs_averaged <- allBetas_DMCs[,45:50]
allBetas_DMCs <- allBetas_DMCs[,1:44]

annotation_col=data.frame(Strain=allPheno$Strain, Sex=allPheno$Sex, CellType=allPheno$CellType,
                          row.names = allPheno$ID)
annotation_col_average=data.frame(CellType=c("B","CD4","CD8","Monocyte","NK","PMN"),
                                  row.names = c("B","CD4","CD8","Monocyte","NK","PMN"))

ann_colors = list(
  Sex = c(Female = "#56c7b4", Male = "#4489b3"),
  CellType = c(B="#1B9E77", CD4="#D95F02",
               CD8 ="#7570B3", Monocyte="#E7298A",
               NK="#E6AB02", PMN="#66A61E"),
  Strain = c(`BALB/c` = "#a83240", `C57BL/6` = "#fab22d", `C57BL/6J` = "#dfed5c")
)

ann_colors_average = list(
  CellType = c(B="#1B9E77", CD4="#D95F02",
               CD8 ="#7570B3", Monocyte="#E7298A",
               NK="#E6AB02", PMN="#66A61E")
)

# Clustered heatmap: CpG probes v. beta values for all samples.
pheatmap(allBetas_DMCs, 
         color = colorRampPalette(c("yellow", "black", "blue"), space = "Lab")(128), 
         annotation_colors = ann_colors,
         annotation_col = annotation_col,
         fontsize = 16,
         main = "Mouse Reference Library",show_rownames = FALSE,
         cluster_cols = T, cluster_rows = T, labels_col = allPheno$CellType, 
         #filename = "~/Desktop/Christensen Lab/Mouse Methylation/Supplementary Heat Map.png", 
         width = 16, height = 14)

# Clustered heatmap: CpG probes v. averaged beta values per cell type.
pheatmap(allBetas_DMCs_averaged, 
         color = colorRampPalette(c("yellow", "black", "blue"), space = "Lab")(128), 
         annotation_colors = ann_colors_average,
         annotation_col = annotation_col_average,
         fontsize = 16,
         main = "Mouse Reference Library",show_rownames = FALSE,
         cluster_cols = T, cluster_rows = T, labels_col = c("CD4","B","CD8","Monocyte","NK","PMN"), 
         #filename = "~/Desktop/Christensen Lab/Mouse Methylation/Final Heat Map.png", 
         width = 16, height = 14)
```

Creating the function to plot model performance for each cell type.

```{r}
cellTypes = c("B", "CD4", "CD8",  
              "Monocyte", "NK", 
              "PMN")

groups<-c( "B"="#1B9E77", 
           CD4="#D95F02",
           CD8 ="#7570B3",
           Monocyte="#E7298A",
           NK="#E6AB02", PMN="#66A61E")

cellcolors = groups #brewer.pal(6,"Set2")
K = length(cellTypes)  # number of cell types

# Plotting function.
combPlotFunction <- function(predDF, obsDF, plotname){
  for(k in 1:K) {
    if (k < 3){
      pred = as.numeric(predDF[,cellTypes[k]])
      obs = as.numeric(obsDF[,cellTypes[k]])
      R2 = cor(obs, pred, method = "pearson")^2
      RMSE = sqrt(sum((pred - obs)^2/length(pred)))
      
      rangexy = range(c(pred,obs))
      
      p <- ggplot(data = data.frame(obs = obs, pred = pred), aes(x = obs, y = pred)) +
        geom_point(size = 5, colour = cellcolors[cellTypes[k]]) + xlim(rangexy) + ylim(rangexy) +
        labs(title = cellTypes[k],
             subtitle = paste("RMSE = ", round(RMSE, 2), "; R2 = ", round(R2, 2), sep = ""),
             x = "Observed (%)",
             y = "Predicted (%)") +
        geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
        theme(plot.title = element_text(face="bold", size=18, hjust = 0.5),
              plot.subtitle = element_text(face="bold", size=16, hjust = 0.5),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 14),
              panel.border = element_rect(color = "black",
                                          fill = NA,
                                          size = 1),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank())
      
      # Regression Line
      reg<-lm(formula = pred ~ obs,data=data.frame(obs = obs, pred = pred)) 
      coeff<-coefficients(reg)          
      intercept<-coeff[1]
      slope<- coeff[2]
      
      p <- p + geom_abline(intercept = intercept, slope = slope, color=cellcolors[cellTypes[k]], size=1)
      
      eval(call("<<-", as.name(paste(plotname, cellTypes[k], sep = "")), p))
      
    } else {
      pred = as.numeric(predDF[,cellTypes[k]])
      obs = as.numeric(obsDF[,cellTypes[k]])
      R2 = cor(obs, pred, method = "pearson")^2
      RMSE = sqrt(sum((pred - obs)^2/length(pred)))
      
      
      p <- ggplot(data = data.frame(obs = obs, pred = pred), aes(x = obs, y = pred)) +
        geom_point(size = 5, colour = cellcolors[cellTypes[k]]) + xlim(c(0,20)) + ylim(c(0,20)) +
        labs(title = cellTypes[k],
             subtitle = paste("RMSE = ", round(RMSE, 2), "; R2 = ", round(R2, 2), sep = ""),
             x = "Observed (%)",
             y = "Predicted (%)") +
        geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
        theme(plot.title = element_text(face="bold", size=18, hjust = 0.5),
              plot.subtitle = element_text(face="bold", size=16, hjust = 0.5),
              axis.title = element_text(size = 14),
              axis.text = element_text(size = 14),
              panel.border = element_rect(color = "black",
                                          fill = NA,
                                          size = 1),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank())
      
      # Regression Line
      reg<-lm(formula = pred ~ obs,data=data.frame(obs = obs, pred = pred)) 
      coeff<-coefficients(reg)          
      intercept<-coeff[1]
      slope<- coeff[2]
      
      p <- p + geom_abline(intercept = intercept, slope = slope, color=cellcolors[cellTypes[k]], size=1)
      
      eval(call("<<-", as.name(paste(plotname, cellTypes[k], sep = "")), p))
    }
  }
}
```

Plotting model performance across cell types for the training and testing data, as well as comparing it to the performance of a Meffil-based model.

```{r}
# Loading the training and testing data.
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/trainingCovariates.RData")
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/trainingBetas.RData")
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/testingCovariates.RData")
load("~/Desktop/Christensen Lab/Mouse Methylation/Data/testingBetas.RData")

# ================================================================================================
# TRAINING 
# ================================================================================================

# Subset the training data to consist of only the CpG sites contained in the IDOL optimized library.
trainingBetas.sub = trainingBetas[IDOLopt_mouse_minfiNoob_300[["IDOL Optimized Library"]],]

# Extract the coefEsts object from the IDOL optimized model.
coefEsts = IDOLopt_mouse_minfiNoob_300[["IDOL Optimized CoefEsts"]]

# Predict cell type composistion.
predictedCellFractionTraining = round(projectWBCnew(trainingBetas.sub, coefEsts)*100,1)

# Visualize the results.
combPlotFunction(predDF = predictedCellFractionTraining, obsDF = trainingCovariates, plotname = "TrainingPlot_")

row1 <- plot_grid(TrainingPlot_B, TrainingPlot_CD4, TrainingPlot_CD8, ncol = 3)
row2 <- plot_grid(TrainingPlot_Monocyte, TrainingPlot_NK, TrainingPlot_PMN, ncol = 3)
TrainingCombPlot <- plot_grid(row1, row2, nrow = 2) + labs(title = "IDOL")

TrainingCombPlot



# ================================================================================================
#TESTING
# ================================================================================================

# Subset the testing data to consist of only the CpG sites contained in the IDOL optimized library.
testingBetas.sub = testingBetas[IDOLopt_mouse_minfiNoob_300[["IDOL Optimized Library"]],]

# Extract the coefEsts object from the IDOL optimized model.
coefEsts = IDOLopt_mouse_minfiNoob_300[["IDOL Optimized CoefEsts"]]

# Predict cell type composistion.
predictedCellFractionTesting = round(projectWBCnew(testingBetas.sub, coefEsts)*100,1)

# Visualize the results.
combPlotFunction(predDF = predictedCellFractionTesting, obsDF = testingCovariates, plotname = "TestingPlot_")

row1 <- plot_grid(TestingPlot_B, TestingPlot_CD4, TestingPlot_CD8, ncol = 3)
row2 <- plot_grid(TestingPlot_Monocyte, TestingPlot_NK, TestingPlot_PMN, ncol = 3)
testingCombPlot <- plot_grid(row1, row2, nrow = 2) + labs(title = "IDOL")

testingCombPlot



# ================================================================================================
# Meffil
# ================================================================================================

# Identifying differentially methylated CpG sites using Meffil's meffil.cell.type.specific.methylation.
diff.DMC.all<-meffil.cell.type.specific.methylation(
  allBetas,
  cell.types=allPheno$CellType,
  number.sites = 50,
  verbose = F
)

# Subset the testing data to consist of only the CpG sites contained in the Meffil identified library.
testingBetas.sub = testingBetas[rownames(diff.DMC.all),]

# Extract the coefEsts object from the Meffil identified model.
coefEsts = diff.DMC.all

# Predict cell type composistion.
predictedCellFractionMeffil = round(projectWBCnew(testingBetas.sub, coefEsts)*100,1)

# Visualize the results.
combPlotFunction(predDF = predictedCellFractionMeffil, obsDF = testingCovariates, plotname = "MeffilPlot_")

row1 <- plot_grid(MeffilPlot_B, MeffilPlot_CD4, MeffilPlot_CD8, ncol = 3)
row2 <- plot_grid(MeffilPlot_Monocyte, MeffilPlot_NK, MeffilPlot_PMN, ncol = 3)
meffilCombPlot <- plot_grid(row1, row2, nrow = 2)

meffilCombPlot

no_stacks <- plot_grid(TrainingCombPlot,testingCombPlot, nrow = 2)
no_stacks
ggsave2(filename = "~/Desktop/Christensen Lab/Mouse Methylation/noStacks300.png", plot = no_stacks, units = "in", height = 15, width = 10)

# ================================================================================================
# Stacked Bar Charts
# ================================================================================================

# Stacked bar chart function.
stackedBarChart <- function(data, title, xlab){
  
  data$Samples <- seq(1,nrow(data))
  data$Samples <- as.factor(data$Samples)
  dfCorrectedForPlot <- data %>% pivot_longer(values_to = "CellTypeProportion",
                                              names_to = "CellType", "B":"PMN", 
                                              values_drop_na = TRUE)
  
  p <- ggplot(data = dfCorrectedForPlot, aes(fill = CellType, x = Samples, y = CellTypeProportion)) +
    geom_bar(position = "stack", stat = "identity", color = "black") + labs(x = xlab,
                                                                            y = "Cell Type Proportion", title = title)
  p <- p + theme(axis.line.x.bottom = element_blank(),
                 axis.line.y.left = element_line(color = 'black'),
                 axis.line.y.right = element_blank(),
                 axis.text.y.right = element_blank(),
                 axis.ticks.y.right = element_blank(),
                 axis.ticks.x = element_blank(),
                 axis.text = element_text(size = 14, color = "black"),
                 axis.title = element_text(size = 14, color = "black"),
                 panel.border = element_blank(),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.background = element_blank(),
                 legend.position = "none",
                 plot.title = element_text(face="bold", size = 24))
  p <- p + scale_fill_manual(values=c(B="#1B9E77", CD4="#D95F02",
                                      CD8 ="#7570B3", Monocyte="#E7298A",
                                      NK="#E6AB02", PMN="#66A61E"))
}

# Stacked bar chart of the cell type proportions across samples using the training samples and IDOL model.
trainStackedBar <- stackedBarChart(as.data.frame(predictedCellFractionTraining), xlab = "Training Samples", title = "(a) Training")
trainStackedBar

# Stacked bar chart of the cell type proportions across samples using the testing samples and IDOL model.
testStackedBar <- stackedBarChart(as.data.frame(predictedCellFractionTesting), xlab = "Testing Samples", title = "(b) Testing")
testStackedBar

# Stacked bar chart of the cell type proportions across samples using the testing samples and Meffil model.
meffilStackedBar <- stackedBarChart(as.data.frame(predictedCellFractionMeffil), xlab = "Testing Samples", title = "(c) Meffil")
meffilStackedBar


# Full Combined Plot ------------------------------------------------------------------------------------
row1 <- plot_grid(trainStackedBar, TrainingCombPlot, ncol = 2, rel_widths = c(2,3))
row2 <- plot_grid(testStackedBar, testingCombPlot, ncol = 2, rel_widths = c(2,3))
row3 <- plot_grid(meffilStackedBar, meffilCombPlot, ncol = 2, rel_widths = c(2,3))
fullCombPlot <- plot_grid(row1, NULL, row2, NULL, row3, nrow=5, rel_heights = c(1,0.05,1,0.05,1))
fullCombPlot
ggsave2(filename = "~/Desktop/Christensen Lab/Mouse Methylation/fullCombPlot.png", plot = fullCombPlot, units = "in", height = 20, width = 16)
```

Comparing errors in cell type prediction between IDOL and Meffil models.

```{r}
# Absolute error by cell type

# IDOL
testingBetas.sub = testingBetas[IDOLopt_mouse_minfiNoob_300[["IDOL Optimized Library"]],]
coefEsts = IDOLopt_mouse_minfiNoob_300[["IDOL Optimized CoefEsts"]]
# Predicting the cell type proportions using the IDOL model.
predictedCellFraction = round(projectWBCnew(testingBetas.sub, coefEsts)*100,1)
predictedCellFraction <- as.data.frame(predictedCellFraction)
testingCovariatesCellTypes <- testingCovariates[,c("CD4T","CD8T","NK","B","PMN","Monocyte")]
# Calculating the errors.
ErrorsIDOL <- data.frame(Holder = c(1,2,3,4,5,6,7,8))
ErrorsIDOL$Bcell <- predictedCellFraction$B - testingCovariatesCellTypes$B
ErrorsIDOL$CD4T <- predictedCellFraction$CD4 - testingCovariatesCellTypes$CD4T
ErrorsIDOL$CD8T <- predictedCellFraction$CD8 - testingCovariatesCellTypes$CD8T
ErrorsIDOL$Monocyte <- predictedCellFraction$Monocyte - testingCovariatesCellTypes$Monocyte
ErrorsIDOL$NK <- predictedCellFraction$NK - testingCovariatesCellTypes$NK
ErrorsIDOL$PMN <- predictedCellFraction$PMN - testingCovariatesCellTypes$PMN
ErrorsIDOL$Holder <- NULL
ErrorsIDOL_corrected_for_plots <- ErrorsIDOL %>% pivot_longer(cols = colnames(ErrorsIDOL), names_to = "CellType", values_to = "Error" )

theme_L_border <- function(colour = "black", size = 1, linetype = 1) {
  structure(
    function(x = 0, y = 0, width = 1, height = 1, ...) {
      polylineGrob(
        x=c(x+width, x, x), y=c(y,y,y+height), ..., default.units = "npc",
        gp=gpar(lwd=size, col=colour, lty=linetype),
      )
    },
    class = "theme",
    type = "box",
    call = match.call()
  )
}

# Meffil
testingBetas.sub.Meffil = testingBetas[rownames(diff.DMC.all),]
coefEsts.Meffil = diff.DMC.all
# Predicting the cell type proportions using the Meffil model.
predictedCellFractionMeffil = round(projectWBCnew(testingBetas.sub.Meffil, coefEsts.Meffil)*100,1)
predictedCellFractionMeffil <- as.data.frame(predictedCellFractionMeffil)
# Calculating the errors.
ErrorsMeffil <- data.frame(Holder = c(1,2,3,4,5,6,7,8))
ErrorsMeffil$Bcell <- predictedCellFractionMeffil$B - testingCovariatesCellTypes$B
ErrorsMeffil$CD4T <- predictedCellFractionMeffil$CD4 - testingCovariatesCellTypes$CD4T
ErrorsMeffil$CD8T <- predictedCellFractionMeffil$CD8 - testingCovariatesCellTypes$CD8T
ErrorsMeffil$Monocyte <- predictedCellFractionMeffil$Monocyte - testingCovariatesCellTypes$Monocyte
ErrorsMeffil$NK <- predictedCellFractionMeffil$NK - testingCovariatesCellTypes$NK
ErrorsMeffil$PMN <- predictedCellFractionMeffil$PMN - testingCovariatesCellTypes$PMN
ErrorsMeffil$Holder <- NULL
ErrorsMeffil_corrected_for_plots <- ErrorsMeffil %>% pivot_longer(cols = colnames(ErrorsMeffil), names_to = "CellType", values_to = "Error" )

# Combined Errors Plot
ErrorsMeffil_corrected_for_plots$Algo <- rep("Meffil",nrow(ErrorsMeffil_corrected_for_plots))
ErrorsIDOL_corrected_for_plots$Algo <- rep("IDOL",nrow(ErrorsIDOL_corrected_for_plots))
ErrorsCombined <- rbind(ErrorsIDOL_corrected_for_plots,ErrorsMeffil_corrected_for_plots)

ErrorsCombinedPlot <- ggplot(data = ErrorsCombined, aes(x = Algo, y = Error, 
                                                        fill = CellType, color = Algo)) +
  geom_boxplot() + scale_fill_manual(values = c(Bcell="#1B9E77", CD4T="#D95F02",
                                                CD8T ="#7570B3", Monocyte="#E7298A",
                                                NK="#E6AB02", PMN="#66A61E")) + 
  scale_color_manual(values = c(IDOL="black", Meffil="blue")) + 
  coord_flip() + labs(x= "Cell Type", y= "Estimate (%) - True (%)",) +
  ylim(-11,11) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = -5, linetype = "dashed", color = "black") + 
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  theme(axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left = element_line(color = 'black'),
        axis.line.y.right = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        panel.border = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.text.y = element_text(colour = c("black","blue")))
ErrorsCombinedPlot

# Global Error Combined Plot
GlobalErrorsCombined <- rbind(ErrorsMeffil_corrected_for_plots[,c(2,3)],ErrorsIDOL_corrected_for_plots[,c(2,3)])

GlobalErrorsCombinedPlot <- ggplot(data = GlobalErrorsCombined, aes(x = Algo, y = Error, color = Algo)) +
  geom_boxplot(width = 0.3) +
  scale_color_manual(values = c(IDOL="black", Meffil="blue")) + 
  coord_flip() + labs(x= "Cell Type", y= "Estimate (%) - True (%)",) +
  ylim(-11,11) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = -5, linetype = "dashed", color = "black") + 
  geom_hline(yintercept = 5, linetype = "dashed", color = "black") +
  theme(axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left = element_line(color = 'black'),
        axis.line.y.right = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.text.y = element_text(colour = c("black","blue")))
GlobalErrorsCombinedPlot

# Total Error Combine Plot
ErrorsIDOL$`Total Error` <- rowSums(abs(ErrorsIDOL))
ErrorsIDOL$Algo <- rep("IDOL", nrow(ErrorsIDOL))
ErrorsMeffil$`Total Error` <- rowSums(abs(ErrorsMeffil))
ErrorsMeffil$Algo <- rep("Meffil", nrow(ErrorsMeffil))

TotalErrorsCombined <- rbind(ErrorsIDOL,ErrorsMeffil)

TotalErrorsCombinedPlot <- ggplot(data = TotalErrorsCombined, aes(x = Algo, y = `Total Error`, color = Algo)) +
  geom_boxplot(width = 0.3) +
  scale_color_manual(values = c(IDOL="black", Meffil="blue")) + 
  coord_flip() + labs(x= "Cell Type", y= "Total Error",) +
  theme(axis.line.x.bottom = element_line(color = 'black'),
        axis.line.y.left = element_line(color = 'black'),
        axis.line.y.right = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.text.y = element_text(colour = c("black","blue")))
TotalErrorsCombinedPlot

col2 <- plot_grid(GlobalErrorsCombinedPlot, TotalErrorsCombinedPlot, nrow = 2, labels = c("B", "C"), label_size = 18, label_fontface = "bold")
col1 <- ErrorsCombinedPlot

finalErrorsPlot <- plot_grid(col1, col2, label_size = 18, label_fontface = "bold", labels = c("A","",""))
finalErrorsPlot

ggsave(filename = "~/Desktop/Christensen Lab/Mouse Methylation/Final Figures/Final Errors Plot.png", plot = finalErrorsPlot, units = "in", height = 12, width = 8)
```

Building a plot to show the most differentially methylated CpG's in each cell type.

```{r}
# Cell type specific CpGs.
cell_specific_CpGs <- c("cg45118317", "cg43964503", "cg43547103", "cg41639691", "cg29054181", "cg34019572")

# Genes that the cell type specific CpGs map to.
genes <- c("CD19","CD4","CD8A","TMEM51OS1","DYRK2","GM31583")

# Beta values for the cell type specific CpGs.
cell_specific_betas <- allBetas[row.names(allBetas) %in% cell_specific_CpGs,]
cell_specific_betas <- t(cell_specific_betas)
cell_specific_betas <- cell_specific_betas[,cell_specific_CpGs]
cell_specific_betas <- as.data.frame(cell_specific_betas)
cell_specific_betas$`Cell Type` <- allPheno$CellType
cell_specific_betas <- cell_specific_betas %>% mutate(`Cell Type` = ifelse(`Cell Type` == "Monocyte","Mono",`Cell Type`))

# Plotting the beta values across cell types for CpG: cg45118317 which maps to CD19.
p1 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,1], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[1], y= "Beta Value",) +
        ggtitle(genes[1]) +
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Plotting the beta values across cell types for CpG: cg43964503 which maps to CD4.
p2 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,2], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[2], y= "Beta Value",) +
        ggtitle(genes[2]) +
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Plotting the beta values across cell types for CpG: cg43547103 which maps to CD8A.
p3 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,3], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[3], y= "Beta Value",) +
        ggtitle(genes[3]) +
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Plotting the beta values across cell types for CpG: cg41639691 which maps to TMEM51OS1.
p4 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,4], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[4], y= "Beta Value",) +
        ggtitle(genes[4]) +
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Plotting the beta values across cell types for CpG: cg29054181 which maps to DYRK2.
p5 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,5], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[5], y= "Beta Value",) +
        ggtitle(genes[5]) +
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Plotting the beta values across cell types for CpG: cg34019572 which maps to GM31583.
p6 <- ggplot(data = cell_specific_betas, 
        aes(x = cell_specific_betas[,7], y = cell_specific_betas[,6], 
            fill = cell_specific_betas[,7])) +
        geom_boxplot() + 
        scale_fill_manual(values = c(B="#1B9E77", CD4="#D95F02",
                                     CD8 ="#7570B3", Monocyte="#E7298A",
                                     NK="#E6AB02", PMN="#66A61E")) + 
        labs(x= cell_specific_CpGs[6], y= "Beta Value",) +
        ggtitle(genes[6]) + 
        theme(axis.line.x.bottom = element_line(color = 'black'),
              axis.line.y.left = element_line(color = 'black'),
              axis.line.y.right = element_blank(),
              axis.text.y.right = element_blank(),
              axis.ticks.y.right = element_blank(),
              panel.border = element_blank(),
              axis.text = element_text(size = 14),
              axis.title = element_text(size = 14),
              plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              legend.position = "none")

# Creating the combined plot.
row1 <- plot_grid(p1, p2, p3, ncol = 3)
row2 <- plot_grid(p4, p5, p6, ncol = 3)
combPlot <- plot_grid(row1, row2, nrow = 2)
combPlot

ggsave(filename = "~/Desktop/Christensen Lab/Mouse Methylation/Final Figures/Cell Type Specific Methylation.png", plot = combPlot, units = "in", width = 15, height = 10)
```
Printing the session info.

```{r}
sessionInfo()
```
